Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-15 prop_iop.tex
prop_letter.tex \section{Gleason} Updating...

\begin{thm}[Empty core for general QIVPMs]\label{thm:Non-extensible-of-Gleason's}In
a Hilbert space of dimension~$d$, there exists a convex QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that $\coreBorn\left(\bar{\mu},\events\right)=\emptyset$.\end{thm}

@@

let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$ and $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$.

@@

defined in Table~\ref{tab:non-Born-QIVPM}, where $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $.

>>>>

\begin{thm}[Empty core for general QIVPMs]\label{thm:Non-extensible-of-Gleason's}In
a Hilbert space of dimension~$d\ge3$, there exists a convex QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that $\coreBorn\left(\bar{\mu},\events\right)=\emptyset$.\end{thm}

@@

let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$,
and $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $.

@@

defined in Table~\ref{tab:non-Born-QIVPM}

### 2017-10-15 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates for the quantum probability measures first, and then introduce
the Born rule which induces a quantum probability measures from each
state~\citep{Born1983,peres1995quantum,544199,Jaeger2007}. Since
the postulates for QIVPMs in Sec.~\ref{sec:Interval-Uncertainty}
are also non-constructive, we want to find a simple rule to construct
a QIVPM as well. When we generalized quantum probability measures
to QIVPMs, we changed the range from the unit interval~$\left[0,1\right]$
to a set of intervals~$\mathscr{I}$. This replacement suggests us
to find a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
which can be used to replace a real-valued probability~$\mu\left(P\right)$
by an interval-valued one~$\iota\left(\mu\left(P\right)\right)$.
In other words, compositing $\iota$ with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
produces an interval-valued function~$\iota\circ\mu:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$
which has the same domain and range as a QIVPM. However, the composited
function~$\iota\circ\mu$ may or may not be a QIVPM. For example,
in a four-dimensional Hilbert space, consider the function 
\begin{equation}
\iota_{1}\left(x\right)=\begin{cases}
\imposs\,, & \textrm{ if }x=0\,;\\
\left[0,\tfrac{1}{2}\right]\,, & \textrm{ if }x\in\left(0,\tfrac{1}{4}\right]\,;\\
\unknown\,, & \textrm{ if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\,;\\
\left[\tfrac{1}{2},1\right]\,, & \textrm{ if }x\in\left[\tfrac{3}{4},1\right)\,;\\
\necess\,, & \textrm{ if }x=1\,.
\end{cases}\label{eq:iota}
\end{equation}
While the function~$\iota_{1}\circ\mu_{\rho_{0}}^{\mathrm{\mathrm{B}}}$
induced by $\rho_{0}=\frac{1}{4}\cdot\proj{0}+\frac{1}{4}\cdot\proj{1}+\frac{1}{4}\cdot\proj{2}+\frac{1}{4}\cdot\proj{3}$
is a QIVPM, the function~$\iota_{1}\circ\mu_{\rho_{1}}^{\mathrm{\mathrm{B}}}$
induced by $\rho_{1}=\frac{1}{8}\cdot\proj{0}+\frac{1}{8}\cdot\proj{1}+\frac{1}{8}\cdot\proj{2}+\frac{5}{8}\cdot\proj{3}$
is \emph{not} a QIVPM~\citep{HOSTunpublished}. In order not to verify
whether $\iota\circ\mu$ is a QIVPM for each individual case, it is
desirable to have some criteria to discriminate different cases. We
will start from the simpler criteria for the dimension~$d\le3$,
and then the general case.

@@

\begin{thm}[Empty core for general QIVPMs]\label{thm:Non-extensible-of-Gleason's}In
a Hilbert space of dimension~$d$, there exists a convex QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that $\coreBorn\left(\bar{\mu},\events\right)=\emptyset$.\end{thm}

To prove the theorem, we first need to construct a QIVPM, and then
verify that no states consistent with it on all projectors. Given
an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2},\ldots,\ket{d-1}\right\} $,
let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$ and $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$.
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly~\citep{HOSTunpublished},
where
\begin{equation}
\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} \,.\label{eq:3-value-intervals}
\end{equation}
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on the three-dimensional Hilbert space. The projectors mapping to
the same interval are listed in the column labeled by $P$ and separated
by commas.}

@@

In other words, if the measurement intervals are too imprecise, the
results we get might contradict to each other, and exclude any possible
state from being consistent with the measurement results. In contrast
with the QIVPM defined in Table~\ref{tab:non-Born-QIVPM}, if a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{B}}$
is induced by a conventional quantum probability measure~$\mu_{\rho}^{\mathrm{B}}$
and a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ considered
in Sec.~\ref{sec:Born}, there is always a state consistent with
it because $\rho\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
Moreover, the collection of states consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower. To state this result, we introduce a few technical notions.
First, the norm of a pure state~$\ket{\psi}$ is defined as usual
$||\psi||=\sqrt{\ip{\psi}{\psi}}$. Then, the norm of the difference
of two states~$\rho$ and $\rho'$ is just the usual operator norm
\begin{equation}
\left\Vert \rho-\rho'\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert \left(\rho-\rho'\right)\ket{\psi}\right\Vert \,,
\end{equation}
which is also commonly known as $2$-norm or spectral norm~\citep{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.

>>>>

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates for the quantum probability measures first, and then introduce
the Born rule which induces a quantum probability measures from each
state~\citep{Born1983,peres1995quantum,544199,Jaeger2007}. Since
the postulates for QIVPMs in Sec.~\ref{sec:Interval-Uncertainty}
are also non-constructive, we want to find a simple rule to construct
a QIVPM as well.

However, the composited function~$\iota\circ\mu$ may or may not
be a QIVPM. For example, in a four-dimensional Hilbert space, consider
the function 
\begin{equation}
\iota_{1}\left(x\right)=\begin{cases}
\imposs\,, & \textrm{ if }x=0\,;\\
\left[0,\tfrac{1}{2}\right]\,, & \textrm{ if }x\in\left(0,\tfrac{1}{4}\right]\,;\\
\unknown\,, & \textrm{ if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\,;\\
\left[\tfrac{1}{2},1\right]\,, & \textrm{ if }x\in\left[\tfrac{3}{4},1\right)\,;\\
\necess\,, & \textrm{ if }x=1\,.
\end{cases}\label{eq:iota}
\end{equation}
While the function~$\iota_{1}\circ\mu_{\rho_{0}}^{\mathrm{\mathrm{B}}}$
induced by $\rho_{0}=\frac{1}{4}\cdot\proj{0}+\frac{1}{4}\cdot\proj{1}+\frac{1}{4}\cdot\proj{2}+\frac{1}{4}\cdot\proj{3}$
is a QIVPM, the function~$\iota_{1}\circ\mu_{\rho_{1}}^{\mathrm{\mathrm{B}}}$
induced by $\rho_{1}=\frac{1}{8}\cdot\proj{0}+\frac{1}{8}\cdot\proj{1}+\frac{1}{8}\cdot\proj{2}+\frac{5}{8}\cdot\proj{3}$
is \emph{not} a QIVPM~\citep{HOSTunpublished}. In order not to verify
whether $\iota\circ\mu$ is a QIVPM for each individual case, it is
desirable to have some criteria to discriminate different cases. We
will start from the simpler criteria for the dimension~$d\le3$,
and then the general case.

@@

\begin{thm}[Empty core for general QIVPMs]\label{thm:Non-extensible-of-Gleason's}In
a Hilbert space of dimension~$d\ge3$, there exists a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
such that $\coreBorn\left(\bar{\mu},\events\right)=\emptyset$.\end{thm}

To prove the theorem, we first need to construct a QIVPM, and then
verify that no states consistent with it on all projectors. Given
an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2},\ldots,\ket{d-1}\right\} $,
let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$,
and
\begin{equation}
\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} \,.\label{eq:3-value-intervals}
\end{equation}
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly~\citep{HOSTunpublished}.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on the three-dimensional Hilbert space. The projectors mapping to
the same interval are listed in the column labeled by $P$ and separated
by commas.}

@@

In other words, if the measurement intervals are too imprecise, the
results we get might contradict to each other, and exclude any possible
state from being consistent with the measurement results. To address
this problem, we are going to consider a particular family of QIVPMs
induced by quantum probability measures. When we generalized quantum
probability measures to QIVPMs, we changed the range from the unit
interval~$\left[0,1\right]$ to a set of intervals~$\mathscr{I}$.
This replacement suggests us to find a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
which can be used to replace a real-valued probability~$\mu\left(P\right)$
by an interval-valued one~$\iota\left(\mu\left(P\right)\right)$.
In other words, compositing $\iota$ with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
produces an interval-valued function~$\iota\circ\mu:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$
which has the same domain and range as a QIVPM. Although there are
infinitely many different choice of $\iota$, the function~$\iota$
satisfies the following homomorphism-like properties are of particular
interest, because it guarantee $\iota\circ\mu$ is always a QIVPM.

\begin{definition}Assume a collection of intervals $\mathscr{I}$
including $\imposs$ and $\necess$. \end{definition}

In contrast with the QIVPM defined in Table~\ref{tab:non-Born-QIVPM},
if a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{B}}$ is induced by a conventional
quantum probability measure~$\mu_{\rho}^{\mathrm{B}}$ and a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
considered in Sec.~\ref{sec:Born}, there is always a state consistent
with it because $\rho\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
Moreover, the collection of states consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower. To state this result, we introduce a few technical notions.
First, the norm of a pure state~$\ket{\psi}$ is defined as usual
$||\psi||=\sqrt{\ip{\psi}{\psi}}$. Then, the norm of the difference
of two states~$\rho$ and $\rho'$ is just the usual operator norm
\begin{equation}
\left\Vert \rho-\rho'\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert \left(\rho-\rho'\right)\ket{\psi}\right\Vert \,,
\end{equation}
which is also commonly known as $2$-norm or spectral norm~\citep{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
